Software architectures for workflow integration in science gateways

### 2016-04-21 paper.tex
Fixed a few todos

See Figure~\subref*{archi:tight}. The workflow engine is tightly
integrated with the science gateway, which means that it is deployed
on the same machine and potentially shares code, libraries and other
software components with the science gateway. For instance, the
workflow engine might be a portlet in a Liferay portal or a controller
in a Ruby on Rails application. The workflow engine and the science
gateway usually share a database where application, users and other
resources are stored. In this model, task and data control are both
initiated from the science gateway. Interactions \texttt{b} and
\texttt{c$_2$} are initiated from the workflow engine while
\texttt{c$_1$} comes from other parts of the science gateway, for
instance a data management interface. As in any other model, the
installation of new workflows in the science gateway (\texttt{a$_2$})
and infrastructure (\texttt{a$_1$}) is done by an administrator, for
security reasons. This is the model adopted in LONI Pipeline
Environment~\cite{dinov2009efficient} and in the Catania Science
Gateway Framework~\cite{Ardizzone2012} which is used for instance in
the DECIDE gateway instance for
neuroimaging~\cite{ardizzone2012decide} (see specific documentation on
workflows\footnote{\url{http://bit.ly/1oQrzvQ}}). \todo{Add
  description of a real system.\note{Tristan}{The Catania Science
    Gateway Framework would be a good illustration here (workflow
    engine is a Java portlet), but I can't find a suitable figure
    about it.}}

%CIPRES~\cite{miller2010creating} (on which the Neuroscience gateway
%described in~\cite{sivagnanam2015early} is based) % \note{Tristan}{There is no workflow engine in there.}

@@

\begin{table*}
\centering
\rowcolors{1}{white}{gray!25}
\begin{tabular}{ll}
  \textbf{Architecture} & \textbf{System} \\
  \hline
  Tight & \pbox{1.5\columnwidth}{
          Catania Science Gateway Framework~\cite{Ardizzone2012}, 
          DECIDE~\cite{ardizzone2012decide}$^{c,n}$, LONI Pipeline Environment$^n$~\cite{dinov2009efficient}
          }
  \\
  Service & \pbox{1.5\columnwidth}{
            Apache Airvata~\cite{marru2011apache}, e-BioInfra~\cite{shahand2015data}$^{w,n}$, HubZero with Pegasus~\cite{CPE:CPE3257}, MoSGrid~\cite{krüger2014mosgrid}$^w$, System in~\cite{wu2010accelerating}, Vine Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual Imaging Platform~\cite{GLAT-13}$^n$, WS-PGRADE/gUSE framework~\cite{Kacsuk2012}, 
            Science gateways in ~\cite{kacsuk2014science}$^w$
            } \\
  Sub-task & CBRAIN and PSOM~\cite{GLAT-16}$^n$, CBRAIN and FSL$^n$\\
  Pool & SHIWA pool~\cite{ROGE-13}\\
  Nested & SHIWA Simulation Platform (Coarse-Grained Interoperability~\cite{terstyanszky2014enabling})$^w$\todo{others}\\
  Import & SHIWA Simulation Platform (Fine-Grained Interoperability)~\cite{plankensteiner-prodan-etal:2013}$^w$
\end{tabular}
\caption{Classification of science gateway based on the architecture used to integrate workflow engines. $^c$: based on the Catania Science Gateway Framework. $^w$: based on WS-PGRADE/gUSE. $^n$: used for neuroimaging.}
\label{table:system-classification}
\end{table*}

@@

\emph{Specific features} include:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item Meta-workflow (\texttt{O$_1$}): ability to describe
  meta-workflows from existing workflows.  \todo{More details about
    why this is important?} \texttt{O$_1$}=0 when the feature is
  available, \texttt{O$_1$}=1 otherwise.
\item Fine-grained debugging (\texttt{O$_2$}): availability of
  fine-grained debugging information about workflow tasks. When
  such information is available, workflow is seen as a white
  box. Fine-grained information about workflow tasks is required to
  properly troubleshoot workflow executions.  \texttt{O$_2$}=0 when
  such information is easy to access, \texttt{O$_2$}=1
  otherwise. \todo{More details about why this is important?}
\end{itemize}
\texttt{O$_1$} and \texttt{O$_2$} are summed to obtain a total number of missing specific
features in the architecture.

@@

\paragraph{Extensibility} Adding a new type of workflow engine requires
to implement the corresponding workflow service, to modify interaction
\texttt{d}, and to implement interactions \texttt{b} and
\texttt{$c_2$} (\texttt{E$_1$}=4). New engine versions can be added by
updating the workflow service without modifying any interaction
(\texttt{E$_2$}=0).  New workflows are added in the science gateway or
in the workflow engine through interactions \texttt{a$_1$} and \texttt{a$_2$}
(\texttt{E$_3$}=2). Adding a new type of infrastructure requires
updates in interactions \texttt{b}, \texttt{$c_1$} and \texttt{$c_2$}
(\texttt{E$_4$}=3).

@@

\paragraph{Extensibility} Adding a new engine type requires to wrap
the engine in the agent and to update interactions \texttt{b} and
\texttt{c$_2$} (\texttt{E$_1$=3}). Updating the version of an engine
is transparent (\texttt{E$_2$=0}) \todo{explain why, also for service}, and integrating a new workflow is
done through interactions \texttt{a$_1$} and \texttt{a$_2$}
(\texttt{E$_3$=2}). Integrating a new infrastructure requires updates
in interactions \texttt{b}, \texttt{c$_1$} and \texttt{c$_2$}
(\texttt{E$_4$=3}).

>>>>

See Figure~\subref*{archi:tight}. The workflow engine is tightly
integrated with the science gateway, which means that it is deployed
on the same machine and potentially shares code, libraries and other
software components with the science gateway. For instance, the
workflow engine might be a portlet in a Liferay portal or a controller
in a Ruby on Rails application. The workflow engine and the science
gateway usually share a database where application, users and other
resources are stored. In this model, task and data control are both
initiated from the science gateway. Interactions \texttt{b} and
\texttt{c$_2$} are initiated from the workflow engine while
\texttt{c$_1$} comes from other parts of the science gateway, for
instance a data management interface. As in any other model, the
installation of new workflows in the science gateway (\texttt{a$_2$})
and infrastructure (\texttt{a$_1$}) is done by an administrator, for
security reasons. This is the model adopted in LONI Pipeline
Environment~\cite{dinov2009efficient}, in the Distributed application
runtime environment (DARE~\cite{maddineni2012distributed}), and in the
Catania Science Gateway Framework~\cite{Ardizzone2012} which is used
for instance in the DECIDE gateway for 
neuroimaging~\cite{ardizzone2012decide} (see specific documentation on
workflows\footnote{\url{http://bit.ly/1oQrzvQ}}). 
%The Catania Science Gateway Framework would be a good illustration here (workflow engine is a Java portlet), but I can't find a suitable figure about it.
%CIPRES~\cite{miller2010creating} (on which the Neuroscience gateway
%described in~\cite{sivagnanam2015early} is based) % \note{Tristan}{There is no workflow engine in there.}

@@

\begin{table*}
\centering
\rowcolors{1}{white}{gray!25}
\begin{tabular}{ll}
  \textbf{Architecture} & \textbf{System} \\
  \hline
  Tight & \pbox{1.5\columnwidth}{
          Catania Science Gateway Framework~\cite{Ardizzone2012}, 
          Distributed application runtime environment (DARE~\cite{maddineni2012distributed}),
          DECIDE~\cite{ardizzone2012decide}$^{c,n}$, LONI Pipeline Environment$^n$~\cite{dinov2009efficient}
          }
  \\
  Service & \pbox{1.5\columnwidth}{
            Apache Airvata~\cite{marru2011apache}, e-BioInfra~\cite{shahand2015data}$^{w,n}$, HubZero with Pegasus~\cite{CPE:CPE3257}, MoSGrid~\cite{krüger2014mosgrid}$^w$, System in~\cite{wu2010accelerating}, Vine Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual Imaging Platform~\cite{GLAT-13}$^n$, WS-PGRADE/gUSE framework~\cite{Kacsuk2012}, 
            Science gateways in ~\cite{kacsuk2014science}$^w$
            } \\
  Sub-task & CBRAIN and PSOM~\cite{GLAT-16}$^n$, CBRAIN and FSL$^n$\\
  Pool & SHIWA pool~\cite{ROGE-13}\\
  Nested & SHIWA Simulation Platform (Coarse-Grained Interoperability~\cite{terstyanszky2014enabling})$^w$\todo{others}\\
  Import & SHIWA Simulation Platform (Fine-Grained Interoperability)~\cite{plankensteiner-prodan-etal:2013}$^w$
\end{tabular}
\caption{Classification of science gateway based on the architecture used to integrate workflow engines. $^c$: based on the Catania Science Gateway Framework. $^w$: based on WS-PGRADE/gUSE. $^n$: used for neuroimaging.}
\label{table:system-classification}
\end{table*}

@@

\emph{Specific features} include:
\begin{itemize}[leftmargin=0cm,itemindent=0.35cm,itemsep=0cm]
\item Meta-workflow (\texttt{O$_1$}): ability to describe
  meta-workflows from existing workflows. Meta-workflows offer an
  additional level of flexibility to build worfklows from reusable
  components. \texttt{O$_1$}=0 when the feature is available,
  \texttt{O$_1$}=1 otherwise.
\item Fine-grained debugging (\texttt{O$_2$}): availability of
  fine-grained debugging information about workflow tasks. When such
  information is available, workflow is seen as a white
  box. Fine-grained information about workflow tasks is required to
  properly troubleshoot workflow executions.  \texttt{O$_2$}=0 when
  such information is easy to access, \texttt{O$_2$}=1 otherwise.
\end{itemize}
\texttt{O$_1$} and \texttt{O$_2$} are summed to obtain a total number of missing specific
features in the architecture.

@@

\paragraph{Extensibility} Adding a new type of workflow engine
requires to implement the corresponding workflow service, to modify
interaction \texttt{d}, and to implement interactions \texttt{b} and
\texttt{$c_2$} (\texttt{E$_1$}=4). New engine versions can be added by
updating the workflow service without modifying any interaction or
component (\texttt{E$_2$}=0). We consider that updating the engine
version in a workflow service is transparent since the only goal of
this component is to wrap the engine.  New workflows are added in the
science gateway or in the workflow engine through interactions
\texttt{a$_1$} and \texttt{a$_2$} (\texttt{E$_3$}=2). Adding a new
type of infrastructure requires updates in interactions \texttt{b},
\texttt{$c_1$} and \texttt{$c_2$} (\texttt{E$_4$}=3).

@@

\paragraph{Extensibility} Adding a new engine type requires to wrap
the engine in the agent and to update interactions \texttt{b} and
\texttt{c$_2$} (\texttt{E$_1$=3}). Updating the version of an engine
is transparent (\texttt{E$_2$=0}) since it only requires updating the
agent that is a component dedicated to the engine. Integrating a new workflow is
done through interactions \texttt{a$_1$} and \texttt{a$_2$}
(\texttt{E$_3$=2}). Integrating a new infrastructure requires updates
in interactions \texttt{b}, \texttt{c$_1$} and \texttt{c$_2$}
(\texttt{E$_4$=3}).
